---
title: "Source code for: Relations among Phonological Processing Skills and Mathematics in Children: A Meta-analysis"
author: "Jason Lo"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    theme: united
---

```{r environment, message=FALSE, warning=FALSE}
source("helper_functions.R")
library(tidyverse)
library(corrr)
library(metaSEM)
mxOption(NULL, 'Number of Threads', parallel::detectCores() - 2)
df <- csv_to_meta_df("main_data_210108.csv", "correlation_data_210108.csv")
```

# Two-stage meta SEM

- Due to large amount of missing data in the raw correlation matrix, we opted to use two-stage meta-analytic structural equation modeling for the analysis. In the first stage, we use a multivariate approach to combine correlation matrices across all studies. In the second stage, we fit a structural equation model with the pooled correlation (See Cheung & Chan, 2005; Cheung & Hafdahl, 2016 for further details).

*References:*
Cheung, M. W. L., & Chan, W. (2005). Meta-analytic structural equation modeling: A two-stage approach. Psychological Methods, 10(1), 40–64. https://doi.org/10.1037/1082-989X.10.1.40

Cheung, M. W. L., & Hafdahl, A. R. (2016). Special issue on meta-analytic structural equation modeling: introduction from the guest editors. Research Synthesis Methods, 7(2), 112–120. https://doi.org/10.1002/jrsm.1212

## Two latent models {.tabset}
- Phonological skill (pho): a latent variable that is measured by pa, ran, and pm
- Cognitive skill (cog): a latent variable that is measured by nviq, vocab, and ef

### Number knowledge
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'number_knowledge')

# Since not all variable will be used from the raw data, we have to call prep_input
# to create a trimmed data set
data <- prep_input(df, observed_variables)

lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef + vocab
                cog ~~ pho
                number_knowledge ~ cog + pho'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1 <-tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```

Result summary:
- Model fit statistic indicate a good fit, with Chi-square X2(12)=20.017, p=.07, CFI = .996, RMSEA =.006
- High correlation between pho and cog, r=.902, p<.001 (2e-16)
- pho --> number_knowledge was significant, coef =.503, p<.05
- cog --> number_knowledge was not significant, coef =.056, p=.829


### Calculation
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'calculation')
data <- prep_input(df, observed_variables)
lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef + vocab
                cog ~~ pho
                calculation ~ cog + pho'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1<-tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```
Model fit statistic indicate a good fit, with Chi-square X2(12)=24.14, p=.02, CFI = .994, RMSEA =.008
High correlation between pho and cog, r=.90, p<.001
pho --> calculation was not significant, coef =.349, p=.129
cog --> calculation was not significant, coef =.275, p=.242


### Word problem
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'word_problem')
data <- prep_input(df, observed_variables)
lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef + vocab
                cog ~~ pho
                word_problem ~ cog + pho'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1<-tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```
Model fit statistic indicate a good fit, with Chi-square X2(12)=12.89, p=.377, CFI = .9995, RMSEA =.0022
High correlation between pho and cog, r=.90, p<.001
pho --> word problem was not significant, coef=-0.417, p=.414
cog --> word problem was significant, coef=1.197, p=.019


### Math accuracy
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'math_accuracy')
data <- prep_input(df, observed_variables)
lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef + vocab
                cog ~~ pho
                math_accuracy ~ cog + pho'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1 <- tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```
Model fit statistic indicate a good fit, with Chi-square X2(12)=31.42, p=.002, CFI = .992, RMSEA =.010
High correlation between pho and cog, r=.90, p<.001
pho --> math acc was not significant, coef=.193, p=.407
cog --> math acc was not significant, coef=.455, p=.057

### Math fluency
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'math_fluency')
data <- prep_input(df, observed_variables)
lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef + vocab
                cog ~~ pho
                math_fluency ~ cog + pho'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1<-tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```
Model fit statistic indicate a good fit, with Chi-square X2(12)=40.56, p<.001, CFI = .985, RMSEA =.012
High correlation between pho and cog, r=.91, p<.001
pho --> math fluency was not significant, coef=.537, p=.055
cog --> math fluency was not significant, coef=-.070, p=.809

## Three IVs models {.tabset}
- Phonological skill (pho): a latent variable that is measured by pa, ran, and pm
- Cognitive skill (cog): a latent variable that is measured by nviq and ef
- Vocabulary knowledge (vocab): a observed variable that is measured by vocab 
- *Due to small number of observation in vocab and cog, the SE is very high for any coefficient related to these factors*

### Number knowledge
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'number_knowledge')

# Since not all variable will be used from the raw data, we have to call prep_input
# to create a trimmed data set
data <- prep_input(df, observed_variables)

lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef
                cog ~~ pho
                cog ~~ vocab
                pho ~~ vocab
                number_knowledge ~ cog + pho + vocab'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1 <-tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```

Result summary:
- Model fit statistic indicate a good fit, with Chi-square X2(9)=13.54, p=0.1398, CFI = 0.9977, RMSEA = 0.0055
- pho --> number_knowledge was not significant 
- cog --> number_knowledge was not significant
- vocab --> number_knowledge was not significant
- *Due to small n in vocab, the standard error is quite high*


### Calculation
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'calculation')
data <- prep_input(df, observed_variables)
lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef 
                cog ~~ pho
                cog ~~ vocab
                pho ~~ vocab
                calculation ~ cog + pho + vocab'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1<-tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```
Model fit statistic indicate a good fit, with Chi-square X2(9)=21.22, p=.012, CFI = 0.9942, RMSEA = 0.0090
pho --> calculation was not significant
cog --> calculation was not significant
vocab --> calculation was not significant

### Word problem
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'word_problem')
data <- prep_input(df, observed_variables)
lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef
                cog ~~ pho
                cog ~~ vocab
                pho ~~ vocab
                word_problem ~ cog + pho + vocab'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1<-tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```
Model fit statistic indicate a good fit, with Chi-square X2(9)=10.9110, p=0.2819, CFI = 0.9990, RMSEA = 0.0037
pho --> word problem was not significant
cog --> word problem was not significant
vocab --> word problem was not significant


### Math accuracy
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'math_accuracy')
data <- prep_input(df, observed_variables)
lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef
                cog ~~ pho
                cog ~~ vocab
                pho ~~ vocab
                math_accuracy ~ cog + pho + vocab'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1 <- tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```
- cannot fit model properly

### Math fluency
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'math_fluency')
data <- prep_input(df, observed_variables)
lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef
                cog ~~ pho
                cog ~~ vocab
                pho ~~ vocab
                math_fluency ~ cog + pho + vocab'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1<-tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```
Model fit statistic indicate a good fit, with Chi-square X2(9)=33.5944, p<.001, CFI = 0.9872, RMSEA = 0.0127
pho --> math fluency was not significant
cog --> math fluency was not significant
vocab --> math fluency was not significant

