---
title: "Source code for: Relations among Phonological Processing Skills and Mathematics in Children: A Meta-analysis"
author: "Jason Lo"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    theme: united
---

```{r environment, message=FALSE, warning=FALSE}
source("helper_functions.R")
library(tidyverse)
library(corrr)
library(metaSEM)
mxOption(NULL, 'Number of Threads', parallel::detectCores() - 2)

# Parse from source
# df <- csv_to_meta_df("main_data_201224.csv", "correlation_data_210108.csv")

# Load from disk
df <- readRDS("df.rds")
```

# Descriptives
```{r}

all_used_vars <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef',
             'number_knowledge', 'calculation', 'word_problem', 'math_accuracy', 'math_fluency')

all_data <- prep_input(df, all_used_vars)


# Number of study per cell
study_n <- pattern.na(all_data$cor, show.na=F, type="tssem")
study_mean <- mean(study_n[lower.tri(study_n)])

# Number of sample per cell
all_n <- pattern.n(all_data$cor, all_data$n)
all_n_mean <- mean(all_n[lower.tri(all_n)])

print(paste("Mean study: ", study_mean, "Mean sample: ", all_n_mean))

# Mean sample size
n_mean <- mean(all_data$n)

# Harmonic mean sample size
h_mean <- length(all_data$n) / sum(1/all_data$n)

print(paste("Mean sample size:", n_mean, "Harmonic mean:", h_mean))


```

## Construct table 1
```{r}
t1 <- matrix(nrow=11, ncol=11)

# Run stage 1 on overall sample (instead of reporting separate stage 1 in each model for simplicity)
overall_stage1 <-tssem1(all_data$cor, all_data$n, method="REM")
overall_summary <- summary(overall_stage1)
t1[lower.tri(t1)] <- coef(overall_stage1, select="fixed")
t1[upper.tri(t1)] <- overall_summary$I2.values[,"Estimate"]
round(t1, 3)
```

# Two-stage meta SEM

- Due to large amount of missing data in the raw correlation matrix, we opted to use two-stage meta-analytic structural equation modeling for the analysis. In the first stage, we use a multivariate approach to combine correlation matrices across all studies. In the second stage, we fit a structural equation model with the pooled correlation (See Cheung & Chan, 2005; Cheung & Hafdahl, 2016 for further details).

*References:*
Cheung, M. W. L., & Chan, W. (2005). Meta-analytic structural equation modeling: A two-stage approach. Psychological Methods, 10(1), 40–64. https://doi.org/10.1037/1082-989X.10.1.40

Cheung, M. W. L., & Hafdahl, A. R. (2016). Special issue on meta-analytic structural equation modeling: introduction from the guest editors. Research Synthesis Methods, 7(2), 112–120. https://doi.org/10.1002/jrsm.1212

## Two latent models
- Phonological skill (pho): a latent variable that is measured by pa, ran, and pm
- Cognitive skill (cog): a latent variable that is measured by nviq, vocab, and ef

### Number knowledge
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'number_knowledge')

# Since not all variable will be used from the raw data, we have to call prep_input
# to create a trimmed data set
data <- prep_input(df, observed_variables)
pattern.na(data$cor, show.na=F, type="tssem")
lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef + vocab
                cog ~~ pho
                number_knowledge ~ cog + pho'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1 <-tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)

plot(stage2)
```

Result summary:
- Model fit statistic indicate a good fit, with Chi-square X2(12)=20.018, p=.07, CFI = .996, RMSEA =.006
- High correlation between pho and cog, r=.902, p<.001 (2e-16)
- pho --> number_knowledge was significant, coef =.503, p<.05
- cog --> number_knowledge was not significant, coef =.056, p=.829


### Calculation
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'calculation')
data <- prep_input(df, observed_variables)
lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef + vocab
                cog ~~ pho
                calculation ~ cog + pho'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1<-tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```
Model fit statistic indicate a good fit, with Chi-square X2(12)=24.14, p=.02, CFI = .994, RMSEA =.008
High correlation between pho and cog, r=.90, p<.001
pho --> calculation was not significant, coef =.349, p=.129
cog --> calculation was not significant, coef =.275, p=.242


### Word problem
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'word_problem')
data <- prep_input(df, observed_variables)
lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef + vocab
                cog ~~ pho
                word_problem ~ cog + pho'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1<-tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```
Model fit statistic indicate a good fit, with Chi-square X2(12)=12.89, p=.377, CFI = .9995, RMSEA =.0022
High correlation between pho and cog, r=.90, p<.001
pho --> word problem was not significant, coef=-0.417, p=.414
cog --> word problem was significant, coef=1.197, p=.019


### Math accuracy
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'math_accuracy')
data <- prep_input(df, observed_variables)
lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef + vocab
                cog ~~ pho
                math_accuracy ~ cog + pho'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1 <- tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```
Model fit statistic indicate a good fit, with Chi-square X2(12)=31.42, p=.002, CFI = .992, RMSEA =.010
High correlation between pho and cog, r=.90, p<.001
pho --> math acc was not significant, coef=.193, p=.407
cog --> math acc was not significant, coef=.455, p=.057

### Math fluency
```{r echo=TRUE}
observed_variables <- c('pa', 'ran', 'pm', 'nviq', 'vocab', 'ef', 'math_fluency')
data <- prep_input(df, observed_variables)
lavaan_spec <- 'pho =~ pa + ran + pm
                cog =~ nviq + ef + vocab
                cog ~~ pho
                math_fluency ~ cog + pho'

plot(lavaan_spec)
RAM_spec <- lavaan2RAM(lavaan_spec, obs.variables=observed_variables)

stage1<-tssem1(data$cor, data$n, method="REM") %>%
      rerun(autofixtau2 = TRUE)

stage2 <- tssem2(stage1, RAM=RAM_spec, intervals.type="z") %>% 
  rerun()

summary(stage2)
plot(stage2)
```
Model fit statistic indicate a good fit, with Chi-square X2(12)=40.56, p<.001, CFI = .985, RMSEA =.012
High correlation between pho and cog, r=.91, p<.001
pho --> math fluency was not significant, coef=.537, p=.055
cog --> math fluency was not significant, coef=-.070, p=.809
