---
title: "Source code for: Relations among Phonological Processing Skills and Mathematics in Children: A Meta-analysis"
author: "Jason Lo"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    theme: united
---

# Important notes
- In the paper, we have a last minute change in table numbering, table 1-5 in the script is table 2-6 in the paper. 

```{r environment, message=FALSE, warning=FALSE}
source("helper_functions.R")

library(multcomp)
library(metafor)
library(tidyverse)


dat <- read.csv('main_data_201224.csv')
dat <- escalc("ZCOR", ri = r, ni = n, data = dat, slab = study)
dat_no_outlier <- readRDS("dat_no_outlier.rds")
df <- readRDS("df.rds")
```


# Egger's test for publication bias
Reference for Egger test using rma.mv(): https://dylancraven.github.io/MetaAnalysis_Course/
- Basically, just add indicator in mod is equivalent to Egger's test
- Indicator can be: vi, standard error: sqrt(vi)... see regtest() for details see https://stats.stackexchange.com/questions/155693/metafor-package-bias-and-sensitivity-diagnostics
```{r}
egger_mv <- rma.mv(yi, vi, mod = I(sqrt(vi)), random = ~ 1 | study, data = dat, slab = study)
summary(egger_mv)
```
To test for publication bias, sampling standard error was used as the indicator in the Egger's regression test, result indicates the publication bias is not significant, z=-1.0591, p=0.2895



# Detailed output dump
## Table 1
```{r}
# Subsetting data in each categorization
dat_math11 <- dat_no_outlier %>%  filter(math1 == 1)  # Basic number knowledge
dat_math12 <- dat_no_outlier %>%  filter(math1 == 2)  # Calculation
dat_math13 <- dat_no_outlier %>%  filter(math1 == 3)  # Word problem
dat_math21 <- dat_no_outlier %>%  filter(math2 == 1)  # Accuracy
dat_math22 <- dat_no_outlier %>%  filter(math2 == 2)  # Fluency
dat_pp1 <- dat_no_outlier %>% filter(PP == 1)  # PA
dat_pp2 <- dat_no_outlier %>% filter(PP == 2)  # RAN
dat_dran <- get_long_cor(df, 'dran')
dat_ndran <- get_long_cor(df, 'ndran')
dat_pp3 <- dat_no_outlier %>% filter(PP == 3)  # PM
dat_dpm <- get_long_cor(df, 'dpm')  
dat_ndpm <- get_long_cor(df, 'ndpm')
dat_sample1 <- dat_no_outlier %>%  filter(sample == 1)  # TD
dat_sample2 <- dat_no_outlier %>%  filter(sample == 2)  # Atypical
dat_sample4 <- dat_no_outlier %>% filter(sample == 4)  # other diseases
dat_grade1 <- dat_no_outlier %>%  filter(grade == 1)  # K1-K3
dat_grade2 <- dat_no_outlier %>%  filter(grade == 2)  # G1-G3
dat_grade3 <- dat_no_outlier %>%  filter(grade == 3)  # G4-G6
dat_design1 <- dat_no_outlier %>%  filter(design == 1)  # Concurrent
dat_design2 <- dat_no_outlier %>%  filter(design == 2)  # Longitudinal


df_list_table1 <- list(dat_no_outlier, dat_math11, dat_math12, dat_math13,
                       dat_math21, dat_math22, dat_pp1, dat_pp2, dat_dran, 
                       dat_ndran, dat_pp3, dat_dpm, dat_ndpm, 
                       dat_sample1, dat_sample2, dat_sample4,
                       dat_grade1, dat_grade2,
                       dat_grade3, dat_design1, dat_design2)

results1 <- lapply(df_list_table1, extract_table1_stat) %>% bind_rows()

name_list_table1 <- c("Overall",
                      "Math1: Basic number knowledge",
                      "Math1:Calculation", 
                      "Math1:Word problem",
                      "Math2: Accuracy", 
                      "Math2: Fluency",
                      "PP: PA", 
                      "PP: RAN", 
                      "PP (New): DRAN",
                      "PP (New): NDRAN",
                      "PP: PM",
                      "PP (New): DPM",
                      "PP (New): NDPM",
                      "Sample: TD", "Sample: Atypical", "Sample: Other",
                      "Grade: K1-K3", "Grade: G1-G3", "Grade: G4-G6",
                      "Design: Concurrent", "Design: Longitudinal")

results1$name <- name_list_table1
results1 <- results1 %>% dplyr::select(name, n_study, k, r, lb, ub, s2)
write.csv(results1, "tables/table1_210113.csv")
```

## Table 2
```{r}
# Create pairwise dataframes
datpair_math1_12 <- rbind(dat_math11, dat_math12) %>% mutate(mod_var = math1)
datpair_math1_23 <- rbind(dat_math12, dat_math13) %>% mutate(mod_var = math1)
datpair_math1_13 <- rbind(dat_math11, dat_math13) %>% mutate(mod_var = math1)
datpair_math2 <- rbind(dat_math21, dat_math22) %>% mutate(mod_var = math2)
datpair_pp_12 <- rbind(dat_pp1, dat_pp2) %>% mutate(mod_var = PP)
datpair_pp_23 <- rbind(dat_pp2, dat_pp3) %>% mutate(mod_var = PP)
datpair_pp_13 <- rbind(dat_pp1, dat_pp3) %>% mutate(mod_var = PP)
datpair_ran <- rbind(dat_dran, dat_ndran) %>% mutate(mod_var = x)
datpair_pm <- rbind(dat_dpm, dat_ndpm) %>% mutate(mod_var = x)
datpair_sample_12 <- rbind(dat_sample1, dat_sample2) %>% mutate(mod_var = sample)
datpair_sample_14 <- rbind(dat_sample1, dat_sample4) %>% mutate(mod_var = sample)
datpair_sample_24 <- rbind(dat_sample2, dat_sample4) %>% mutate(mod_var = sample)

datpair_grade_12 <- rbind(dat_grade1, dat_grade2) %>% mutate(mod_var = grade)
datpair_grade_23 <- rbind(dat_grade2, dat_grade3) %>% mutate(mod_var = grade)
datpair_grade_13 <- rbind(dat_grade1, dat_grade3) %>% mutate(mod_var = grade)
datpair_design <- rbind(dat_design1, dat_design2) %>% mutate(mod_var = design)

# Bundle into a list (only for categorical moderator)
df_list_table2 <- list(datpair_math1_12, datpair_math1_23, datpair_math1_13,
                       datpair_math2, datpair_pp_12, datpair_pp_23, datpair_pp_13,
                       datpair_ran, datpair_pm,
                       datpair_sample_12, datpair_sample_14, datpair_sample_24, 
                       datpair_grade_12, datpair_grade_23, datpair_grade_13,
                       datpair_design)


# Run stats
results2 <- lapply(df_list_table2, extract_table_stat_categorical_mod) %>% bind_rows()
results2 <- rbind(results2, extract_table_stat_continous_mod(dat_no_outlier))


# Give each row of analysis a name
name_list_table2 <- c("Math1: Basic vs. Calculation", 
                      "Math1: Calculation vs. Word",
                      "Math1: Basic vs. Word",
                      "Math2: Accuracy vs. Fluency",
                      "PP: PA vs. RAN",
                      "PP: RAN vs. PM",
                      "PP: PA vs. PM",
                      "RAN: Digit vs. Non-digit",
                      "PM: Digit vs. Non-digit",
                      "Sample: TD vs. Atypical",
                      "Sample: TD vs. Other",
                      "Sample: Atypical vs. Other",
                      "Grade: K1-K3 vs. G1-G3",
                      "Grade: G1-G3 vs. G4-G6",
                      "Grade: K1-K3 vs. G4-G6",
                      "Design: Concurrent vs. Longitudinal",
                      "Age") # Assume the last


results2$name <- name_list_table2

# Adjusting p-value for entire table using false discovery rate
results2$adj_p_fdr <- p.adjust(results2$p, method = "fdr")

results2 <- results2 %>% dplyr::select(name, n_study, k, beta, se, lb, ub, z, p, adj_p_fdr)
write.csv(results2, "tables/table2_210113.csv")

```


## Prep for table 3-5
```{r}
df_list_master_table345 <- list(datpair_math1_12, datpair_math1_23, datpair_math1_13,
                             datpair_math2, 
                             datpair_sample_12, datpair_sample_14, datpair_sample_24, 
                             datpair_grade_12, datpair_grade_23, datpair_grade_13,
                             datpair_design)

name_list_table345 <- c("Math1: Basic vs. Calculation", 
                 "Math1: Calculation vs. Word",
                 "Math1: Basic vs. Word",
                 "Math2: Accuracy vs. Fluency",
                 "Sample: TD vs. Atypical",
                 "Sample: TD vs. Other",
                 "Sample: Atypical vs. Other",
                 "Grade: K1-K3 vs. G1-G3",
                 "Grade: G1-G3 vs. G4-G6",
                 "Grade: K1-K3 vs. G4-G6",
                 "Design: Concurrent vs. Longitudinal",
                 "Age")

run_table_345 <- function(pp_filter){
  filter_function <- function(df){
    return(df[df$PP==pp_filter,])
  }
  
  # Apply filter
  df_cat_mod_list <- lapply(df_list_master_table345, filter_function)
  df_cont_mod_age <- filter_function(dat_no_outlier)
  
  # Run stats
  results <- lapply(df_cat_mod_list, extract_table_stat_categorical_mod) %>% bind_rows()
  results <- rbind(results, extract_table_stat_continous_mod(df_cont_mod_age))
  results$name <- name_list_table345
  
  # Adjusting p-value for entire table using false discovery rate
  results$adj_p_fdr <- p.adjust(results$p, method = "fdr")
  
  results <- results %>% dplyr::select(name, n_study, k, beta, se, lb, ub, z, p, adj_p_fdr)
  return(results)
}
```


## Table 3-5
```{r}

results3 <- run_table_345(pp_filter = 1)
write.csv(results3, "tables/table3_210113.csv")

results4 <- run_table_345(pp_filter = 2)
write.csv(results3, "tables/table4_210113.csv")

results5 <- run_table_345(pp_filter = 3)
write.csv(results3, "tables/table5_210113.csv")

```

# Formatted output
## Table 1

```{r}
results1
```

## Table 2
```{r}
results2
```

## Table 3
```{r}
results3
```

## Table 4
```{r}
results4
```

## Table 5
```{r}
results5
```